---
title: "Querying RESULTS silviculture data using R"
author: "Hardy Griesbauer"
date: "06/04/2020"
output: 
  
  html_document
 

---

<!--

  Copyright 2019 Province of British Columbia

  Licensed under the Apache License, Version 2.0 (the "License");
  you may not use this file except in compliance with the License.
  You may obtain a copy of the License at

  http://www.apache.org/licenses/LICENSE-2.0

  Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
  WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
  See the License for the specific language governing permissions and limitations under the License.

-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

### bcdata and bcmaps R packages
The `bcdata` and `bcmaps` packages now provide an easy way to query, download and visualize select BC government datasets through the R interface. This short vignette will demonstrate a few ways that we can do this using RESULTS and other datasets.  These packages have been developed by Andy Teucher, Sam Albers, Stephanie Hazlitt and others, and more information can be found here: https://bcgov.github.io/bcdata/

While BC government data can usually be downloaded through your browser at catalogue.data.gov.bc.ca, I think there are some advantages to using the `bcdata` package to download data form within R:

1. You are not limited to download limits;
2. You can query the data before downloading to extract only what you need;
3. After downloading, you can quickly work with the data in R; and
4. Development team is very responsive and great to work with!  

## Getting started
First thing we need to do is make sure we have some packages installed.  If you don't have the package already installed on your machine, run the following code to install from CRAN:

```{r install bcdata,eval=FALSE}
# Install bcdata from CRAN
install.packages("bcdata")

# Install bcmaps from CRAN
install.packages("bcmaps")

```

To start the demo, let's load some other libraries:
```{r,message=FALSE,warning=FALSE}

library(tidyverse)
library(bcdata)
library(mapview)
library(bcmaps)
library(bcmapsdata)
library(here)

```

### Different ways to use the bcdata package

The main functions in `bcdata` are:

1. bcdc_browse() - Open the catalogue in your default browser
2. bcdc_search() - Search records in the catalogue
3. bcdc_search_facets() - List catalogue facet search options
4. bcdc_get_record() - Print a catalogue record
5. bcdc_get_data() - Get catalogue data
6. bcdc_query_geodata() - Get & query catalogue geospatial data available through a Web Service

In general, I use an iterative workflow with these packages:

1. Search for datasets in the catalogue;
2. Query the dataset to understand the data structure;
3. Filter for records of interest; and
4. Download


```{r search}

bcdc_search("silviculture")

```

### Silviculture data in BC
Silviculture information in BC is largely collected and tracked through the RESULTS database (Reporting Silviculture Updates and Land Status Tracking System).  The database stores silviculture-related information on the management of openings, disturbances, silviculture activities and obligation declarations as required by the Forest and Range Practices Act.  Data in RESULTS are publically available through catalogue.data.gov.bc.ca.  More information on RESULTS can be found here: https://www2.gov.bc.ca/gov/content/industry/forestry/managing-our-forest-resources/silviculture/silviculture-reporting-results

## Application of bcdata and bcmaps to RESULTS data

### Reforestation with climate considerations in the Prince George District
The Omineca Region's climate action plan has targets around reforestation with tree species that will likely remain well-adapted to projected future climates.  Douglas-fir and western larch have been identified as tree species that will remain viable into the future.

*Question: how much western larch is being planted in the Prince George District?*

We can gather information to answer this question using RESULTS - silviculture forest cover dataset.  First, let's take a look at the dataset contents using `bcdc_query_geodata`:

```{r examine RESULTS silv layer}

bcdc_query_geodata("results-forest-cover-silviculture") 

```

The output from this query shows that this dataset has 866,728 features and 159 fields.  Each feature is a treatment unit within a harvested opening, and contains information on the leading five tree species that are present in each treatment unit, including stems per hectare, age, and height.

This dataset would be too large (~1GB) to download efficiently.  So, let's use `filter` to refine our query:

1. Filter openings that are in Prince George District; and
2. Filter treatment units that have Douglas-fir or western larch present.

First, let's filter for openings in the PG District.  We'll use both `bcmaps` and `bcdata` package to do this by downloading a shapefile for the PG District and use that as a bounding box:

```{r filter openings for DPG}

# First, let's take a look at the natural resource district dataset
bcdc_query_geodata("natural-resource-nr-district")
```

We can see that there are 23 NR districts in the province.  District codes are stored in the 'ORG_UNIT' column  Let's filter districts for DPG (Prince George) and download it as a shapefile.

```{r download DPG}

dpg<-  # Create new variable called dpg
  bcdc_query_geodata("natural-resource-nr-district") %>%  # query the nr district dataset 
  filter(ORG_UNIT=="DPG") %>% # filter for Prince George District
  collect() # and download it 

# Alternatively, load from workspace
# load(here("data","dpg.RData"))

```

This is a small file, and shouldn't take long to download.  We can plot it to double check:

```{r plot DPG}

dpg %>% 
  ggplot()+
  geom_sf()

```

Now we have a shapefile that we can use as a bounding box to filter and download records in the RESULTS - silviculture layer.  We'll just download openings that have larch planted, to make for a smaller file size.

*Note: the chunk below can take a minute or so to run.  To save time, I've already run the chunk and saved the output to an .RData file, which we'll load into our workspace in a subsequent chunk*

```{r,eval=FALSE}

# First, let's list what tree species we are interested in
 sppList=c("LW")
 
# Now let's filter the records and download the data

  treesDPG<- # create new variable by
    bcdc_query_geodata("results-forest-cover-silviculture") %>% # querying the results silviculture layer and
   
    filter(WITHIN(dpg)) %>% # filter for records that are within the DPG and
    
    filter(S_SPECIES_CODE_1 %in% sppList| # filter for our species within any of the five species per TU
            S_SPECIES_CODE_2 %in% sppList|
            S_SPECIES_CODE_3 %in% sppList|
            S_SPECIES_CODE_4 %in% sppList|
            S_SPECIES_CODE_5 %in% sppList) %>% 
    
    collect() # and download the file

```

Let's load this dataset into our workspace:
```{r load lw data}

load(here("data","treesDPG.RData"))

```

Now we can plot the openings and see how things look:

```{r map larch plantations in DPG}

mapview(treesDPG,zcol="S_SILV_LABEL",legend=FALSE)+dpg
 

```


### Summaries
We can also create some quick summaries of the data, treating the attribute table as a data frame in R:

#### Question: what is the age distribution of larch plantations in DPG?
```{r,warning=FALSE}

treesDPG %>% 
  st_drop_geometry() %>% # drop geometry
  mutate(age=2020-REFERENCE_YEAR+S_SPECIES_AGE_1) %>% 

  # plotting
  ggplot()+
  aes(x=age,y=FEATURE_AREA_SQM/10000)+
  geom_bar(stat="sum")+
  ylab("Sum treatment unit area (ha)")+
  theme(legend.position="none")+
  scale_x_continuous(name="Stand age", limits=c(2,30),
                     breaks=seq(from=5,to=30,by=5))+
  ggtitle("Area planted with larch by age in DPG, as of 2020")



```

#### Where are the older larch plantations in DPG?
```{r}
treesDPG %>% 
  mutate(age=2020-REFERENCE_YEAR+S_SPECIES_AGE_1+1) %>% # create age column
  filter(age>20) %>% # filter for stand ages older than 20
  mapview(zcol="age") # map

```

#### Question: what is the BEC distribution of larch plantations in DPG?
The `bcmaps` package also contains BEC maps, which can be used to look at larch distribution by BGC unit in DPG.  BEC data are loaded into the workspace using the `bec()` function, and then can be joined to an existing dataset using the `st_join` function.  This takes awhile, so I did it beforehand.

```{r,eval=FALSE}

# First, download BEC data 
 becData<-bec() 

# Now join with dataset
treesDPG<-  
  st_join(treesDPG,bgcPG[,"MAP_LABEL"])

```

Let's map blocks by BGC unit:
```{r}
    
mapview(treesDPG,zcol="MAP_LABEL")

```


#### Summarize area planted with larch by BGC unit

```{r,warning=FALSE}

treesDPG %>% 
  st_drop_geometry() %>% # drop geometry
  group_by(MAP_LABEL) %>% # group polygons by BGC unit
  summarise(Area=sum(FEATURE_AREA_SQM)/10000) %>% 
  # plotting
  ggplot()+
  aes(x=MAP_LABEL,y=Area)+
  geom_bar(stat="Identity")+
  ylab("Sum treatment unit area (ha)")+
  theme(legend.position="none")+
  ggtitle("Area planted with larch by BGC unit in DPG")+
  xlab("BGC unit")



```

---